On the number of tiles touched by a\\ line segment on a rectangular grid

### 2021-08-02 paper.tex
Some discussion about average-number result

The problems studied in this paper are related to \ldots ***maybe a short paragraph here mentioning related problems: Buffon's needle, Buffon-Laplace problem. Some references: ``Buffon's Noodle Problem'', J.F.~Ramaley, 1969. ``On Laplace's Extension of the Buffon Needle Problem'', B.J.~Arnow, 1994. \\
\url{https://www.jstor.org/stable/2317945} \\
\url{https://www.jstor.org/stable/2687085}

@@

\begin{proposition}
Given $a, b, \len \in \mathbb R^+$, consider a grid with parameters $a, b$ and a uniformly random segment of length $\len$, as defined above. Let the random variables $i$, $j$ represent the normalized dimensions of the canonical rectangle. For $n \in \mathbb N$,
\begin{equation}
\label{eq: Pr i geq n}
\Pr[i \geq n] = \begin{cases}
\displaystyle
1 &\text{if\ \ } \displaystyle n =1 \\[1 mm]
\displaystyle
\frac{2\len}{\pi a} \left(g\left(\frac{a(n-1)}{\len}\right)-g\left(\frac{a(n-2)}{\len} \right)\right) &\text{if\ \ } \displaystyle 2 \leq n < \frac \len a + 1 \\[4 mm]
\displaystyle
\frac{2\len}{\pi a}\,\, \left(1 - g\left(\frac{a(n-2)}{\len}\right)\right) &\text{if\ \ } \displaystyle \frac\len a + 1 \leq n < \frac\len a+2 \\[3 mm]
\displaystyle
0 &\text{if\ \ } \displaystyle \frac \len a + 2 \leq  n; \\
\end{cases}
\end{equation}
and $\Pr[j \geq n]$ is obtained from \eqref{eq: Pr i geq n} replacing $a$ by $b$.
\end{proposition}

@@

Consider the event $x_2 \geq a(n-1)$. There are three possibilities depending on $x_1$, $n$ and $\len$. If $a(n-1) < \len$, the length $\len$ is enough for $x_2$ to exceed $a(n-1)$ for some angles $\orient$, regardless of $x_1$. This is depicted in Figure~\ref{fig: Pr_i_full}, where the thicker line on the arc represents those angles for which $x_2 > a(n-1)$, for a specific $x_1$. If $\len \leq a(n-1) < \len+a$, the length is enough only if $x_1 \geq a(n-1)-\len$. This can be seen in Figure~\ref{fig: Pr_i_notfull}, where the thicker line again represents the angles for which this happens, for a specific $x_1 \geq a(n-1)-\len$. Lastly, if $\len+a \leq a(n-1)$ it is impossible that $x_2$ exceeds $a(n-1)$. Note that the value $y_1$ is irrelevant for this.

@@

***Explain that the inverse question is straightforward, because $\funta$ is monotone increasing, thus it has an inverse function (defined on the range of $\funta$). In fact, as will be seen, $\funta$ is an affine function, and therefore it is immediate to obtain its inverse.

***Some mention that there is no need to study square grids or the case of integer-valued lengths separately.

***Relation the problem of the Buffon needle. On one hand, the formula for b=infinity gives the average number of lines crossed in the Buffon experiment. On the other hand, the intermediate values $\Pr[i \geq i_0]$ (proposition) are a generalization of the probability of crossing at least a line in the Buffon experiment, when the needle endpoint is only allowed to move in a smaller vertical strip.

***Mention particular values: a=b=1 gives...; a=b=4/pi gives len+1. Can we find any ``meaning'' for these particular cases?

***The function is affine, not linear. This is related to the fact that the parts of the segment near the endpoints can touch several tiles

>>>>

The problems studied in this paper are related to \ldots ***maybe a short paragraph here mentioning related problems: Buffon's needle, Buffon-Laplace problem. Some references: ``Buffon's Noodle Problem'', J.F.~Ramaley, 1969. ``On Laplace's Extension of the Buffon Needle Problem'', B.J.~Arnow, 1994. \emph{An Introduction to Geometrical Probability}, A.M~Mathai, section 1.1. \\
\url{https://www.jstor.org/stable/2317945} \\
\url{https://www.jstor.org/stable/2687085}

@@

\begin{proposition}
\label{eq: prop Pr i}
Given $a, b, \len \in \mathbb R^+$, consider a grid with parameters $a, b$ and a uniformly random segment of length $\len$, as defined above. Let the random variables $i$, $j$ represent the normalized dimensions of the canonical rectangle. For $n \in \mathbb N$,
\begin{equation}
\label{eq: Pr i geq n}
\Pr[i \geq n] = \begin{cases}
\displaystyle
1 &\text{if\ \ } \displaystyle n =1 \\[1 mm]
\displaystyle
\frac{2\len}{\pi a} \left(g\left(\frac{a(n-1)}{\len}\right)-g\left(\frac{a(n-2)}{\len} \right)\right) &\text{if\ \ } \displaystyle 2 \leq n < \frac \len a + 1 \\[4 mm]
\displaystyle
\frac{2\len}{\pi a}\,\, \left(1 - g\left(\frac{a(n-2)}{\len}\right)\right) &\text{if\ \ } \displaystyle \frac\len a + 1 \leq n < \frac\len a+2 \\[3 mm]
\displaystyle
0 &\text{if\ \ } \displaystyle \frac \len a + 2 \leq  n; \\
\end{cases}
\end{equation}
and $\Pr[j \geq n]$ is obtained from \eqref{eq: Pr i geq n} replacing $a$ by $b$.
\end{proposition}

@@

Consider the event $x_2 \geq a(n-1)$. There are three possibilities depending on $x_1$, $n$ and $\len$. If $a(n-1) < \len$, the length $\len$ is enough for $x_2$ to exceed $a(n-1)$ for some angles $\orient$, regardless of $x_1$. This is depicted in Figure~\ref{fig: Pr_i_full}, where the section of the arc with solid line represents those angles for which $x_2 > a(n-1)$. If $\len \leq a(n-1) < \len+a$, the length is enough only if $x_1 \geq a(n-1)-\len$, which corresponds to the shaded region in Figure~\ref{fig: Pr_i_notfull}. Lastly, if $\len+a \leq a(n-1)$ it is impossible that $x_2$ exceeds $a(n-1)$. Note that the value $y_1$ is irrelevant for this.

@@

In view of Theorem~\ref{theo: funta, form}, the average number of touches tiles has a remarkably simple dependence on the segment length, namely an affine function. Conversely, for any $t>1$ it is immediate to compute the length of a random segment that touches $t$ tiles on average, given as $\funta^{-1}(t)$.

The results in this section make clear the relationship between the problem considered here (average number of squares touched by a random segment of fixes length) and the Buffon's needle problem (probability that a random segment of fixed length crosses a line of a regular structure of parallel strips) \cite[section 1.1]{Mathai99}. Firstly, since $\pi$ is involved in \eqref{eq: funta, form}, it is possible to design a simple probabilistic experiment to estimate the value of $\pi$, as in Buffon's original experiment. For example, choosing $a=b=\len=1$ gives $\funta(\len) = 1+1/\pi$.

Secondly, a grid with $b \rightarrow \infty$ corresponds to Buffon's arrangement of parallel strips with spacing $a$. Thus
\begin{equation}
\label{eq: number of lines crossed, Buffon}
\lim_{b \rightarrow \infty} (\funta(\len)-1) = \frac{2\len}{pi a})
\end{equation}
gives the average number of lines crossed in Buffon's experiment. For $\len \leq a$ the segment can only cross one line, and \eqref{eq: number of lines crossed, Buffon} coincides with the probability of crossing \cite[section 1.1]{Mathai99}.

Lastly, $\Pr[i \geq n]$ as computed in Proposition~\ref{eq: prop Pr i} for $n \geq 2$ can be seen as a generalization of the probability of crossing at least one line in Buffon's experiment, when a needle endpoint is only allowed to move in a narrower region of width $n-1$ times smaller than the width of the strips (instead of moving across the full strip as in Buffon's setting).
